\documentclass[10pt]{beamer}
\mode<presentation>
  {
    \usetheme{CambridgeUS}
    \setbeamercovered{transparent}
  }
\usefonttheme[]{serif}
\usepackage[english,ukrainian]{babel}
\usepackage[OT1,T2A]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{times}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}


\mathchardef\ordinarycolon\mathcode`\:
\mathcode`\:=\string"8000
\begingroup \catcode`\:=\active
  \gdef:{\mathrel{\mathop\ordinarycolon}}
\endgroup
\setlength{\parindent}{1cm} \setlength{\parskip}{1ex plus 0.5ex
minus 0.2ex}

%----------------------------------------------------------------
\title{Trident Genesis}
\subtitle{The Software Development Platform}

\author{FMS}
\institute{Fielden R'n'D}

\date{Lviv -- 2009}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%title%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{\titlepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Making of Software}
\frame{
	\frametitle{Software development with Delphi: Problems I}
	\Large{
	\begin{itemize}
  	\item High level of development complexity -- too much of tech details.
  	\item Poor to none code reuse.
  	\item Referential integrity issues.
	\item Concurrent data modification issues.
	\item No unit tests to ensure business logic is not broken after modification.
	\end{itemize}
      }
}
\frame{
	\frametitle{Software development with Delphi: Problems II}
	\Large{
	\begin{itemize}
	\item Hardcoded values.
	\item Data duplication at the DB level (no surrogate IDs).
	\item Difficult to modify UI layout due to pixel precision positioning.
	\item User interface usability issues (intrusive validation etc).
	\item RDBMS dependency (SP + proprietary features).
	\end{itemize}
      }
}
\frame{
	\frametitle{Browser-based web applications with Java}
	\Large{
	\begin{itemize}
  	\item Browser dependency.
  	\item Lack of UI richness and interactiveness (HTML limitation, session management)
  	\item Complex development model (HTML, CSS, JavaScript, AJAX, Web-framework...) -- huge!
	\item Still too much of technical details.
	\end{itemize}
      }
}
\frame{
	\frametitle{Domain-Driven Design}
	\Large{
	\begin{block}{}
	 Software design approach based on two premises:
	\end{block}
	\begin{itemize}
  	\item Complex domain designs should be based on a business model.
  	\item The primary focus should be on the domain and domain logic -- not a particular technology used to implement the system.
	\end{itemize}
	}
}
\frame{
	\frametitle{Domain-Driven Design: Metaphors}
	\Large{
	\begin{itemize}
  	\item Entities: continuity and identity.
  	\item Value Objects: aggregated data.
  	\item Services: business logic
	\item Repositories: methods for storing domain objects.
	\item Factories: methods for creating domain objects.
	\end{itemize}
	}
}
\section{Trident Genesis Platform}
\frame{
	\frametitle{What is TG?}
	\Large{
	\begin{block}{}
	Trident Genesis (TG) is an application platform for developing rich internet applications where the main emphasis is made on Domain-Driven Design.
	\end{block}
	\begin{block}{}
	The platform tries to take away the technical hurdles as much as possible while providing a complete control to developers (i.e. no code generation, no special tools etc.).
	\end{block}
      }
}
\frame{
	\frametitle{Core Principles I}
	\Large{
	\begin{itemize}
  	\item Domain centric application development with a life cycle revolving around business entities and logic.
  	\item Easy development of business entities by providing common implementation for identity, equality, natural ordering etc.
  	\item Declarative programming.
	\item Highly configurable.
	\item Web-enabled -- supports REST architecture.
	\item Client Application Framework for life cycle management.
	\end{itemize}
      }
}
\frame{
	\frametitle{Core Principles II}
	\Large{
	\begin{itemize}
	\item Ease of user interface (UI) development in a domain centric way.
	\item Scalable architecture -- caters for development of small and large business systems.
	\item RDBMS independent by means of using O/R Mapper.
	\item Based on well-known FOSS Java technologies such as Guice, Hibernate, Swing reducing the learning curve.
	\item Pure Java (no need to know HTML, CSS, JavaScript and other browser-related technologies)
	\end{itemize}
      }
}
\frame{
	\frametitle{Main Advantages}
	\Large{
	\begin{itemize}
	\item Improves maintainability (5-20 times code reduction)
	\item Increases speed of development.
	\item Enables truly Rich Internet Application development.
	\item Covers a complete development life cycle -- from DB to UI.
	\item OS and RDBMS independent.
	\item Ensures data integrity automatically (covers concurrent modifications)
	\item Based on pure Java (single technology -- reduces learning curve)
	\end{itemize}
      }
}
\frame{
	\frametitle{Business Model around TG}
	\Large{
	\begin{itemize}
	\item The problems TG solves are not unique to FMS.
	\item TG is a viable solution for others.
	\item We eat our own dog food -- very important for adoption by others.
	\item Dual licensing (FOSS and commercial) to gain popularity.
	\item Commercial support including training and mentoring.
	\item The audience: existing Java and non-Java developers in ISV firms as well as IT departments for in-house custom solution development.
	\end{itemize}
      }
}
\frame{
	\frametitle{Future directions}
	\Large{
	\begin{itemize}
	\item TG development patterns.
	\item Development tools (Eclipse Plug-ins) for automating creation of business entities and stubs for associated business rules.
	\item Ready to use business solutions for specific business domains (EAM, ERP, CRM).
	\item Domain Specific Languages using Scala.
	\item TG-based SaaS solutions.
	\end{itemize}
      }
}

\section{Other}
\frame{
	\frametitle{FMS.ua Development Infrastructure}
	\Large{
	\begin{itemize}
	\item Eclipse IDE -- used for development of TG and TG-based solutions.
	\item Trac -- requirement and bug tracking management tool.
	\item Hudson -- continuous integration (CI) tool.
	\item Maven -- a software project management and comprehension tool.
	\item Sonar -- code quality management tool.
	\item Subversion -- a version control system.
	\end{itemize}
      }
}

\end{document}